Reformulating phylogenetic mixed models to improve flexibility and speed

### 2023-10-24 phyloglmm_ms_resubmit.tex
latex tweaks

\newcommand{\plmefour}{phyloglmm/lme4}
\newcommand{\pgTMB}{phyloglmm/glmmTMB}
\newcommand{\phyvar}{\sigma} %% phylogenetic var: allow sigma/Sigma switch
\newcommand{\phyv}[1]{\ensuremath{\phyvar^2_{\mathrm{phy}_\mathrm{#1}}}}
\newcommand{\phys}[1]{\ensuremath{\phyvar_{\mathrm{phy}_\mathrm{#1}}}}
\newcommand{\bphys}[1]{\ensuremath{{\bf b}_{\mathrm{phy}_{\mathrm{#1}}}}}
\newcommand{\bsps}[1]{\ensuremath{{\bf b}_{\mathrm{sp}_{\mathrm{#1}}}}}
% \pagenumbering{gobble}
\linenumbers

@@

We extend the simulation model by adding multiple groups where each group has one observation per species. 
The multi-group model is a generalization of multiple-site models used in community ecology to model phylogenetic attraction and repulsion \citep{helmus2007separating}. 
The full multi-group model is as follows: 
\begin{equation}
\begin{aligned}
\bY & = \dist(g^{-1}(\bmu), \phi) \\
\bmu & = (\beta_0 + \bphys{int} + \bsps{int} + {\bf b}_{\mathrm{group}}}) + (\beta_1 + \bphys{slope} + \bsps{slope}) \bm{t} + {\bf b}_{\mathrm{sp:group}}} + \besp \\
(\bphys{int}, \bphys{slope}) & \sim \textrm{MVN} \left( 0, \begin{bmatrix}
\phyv{int} & \phys{int-slope} \\ 
\phys{int-slope} & \phyv{slope}
\end{bmatrix}
\right) \\
(\bf{b_{\mathrm{sp_{int}}}, b_{\mathrm{sp_{slope}}}}) & \sim \textrm{MVN} \left( 0, \begin{bmatrix}
\sigma^2_{\mathrm{sp_{int}}} & \sigma_{\mathrm{sp_{int-slope}}} \\ 
\sigma_{\mathrm{sp_{int-slope}}} & \sigma^2_{\mathrm{sp_{slope}}}
\end{bmatrix}
\right) \\
\bf{b_{\mathrm{group}}} & \sim \textrm{MVN} ( 0 , \sigma^2_{\mathrm{group}}) \\
\bf{b_{\mathrm{sp:group}}} & \sim \textrm{MVN} (0, \mathrm{I}_{\mathrm{group}} \otimes \sigma^2_{\mathrm{phy}}) \\
\besp & \sim \textrm{N} ( 0 , \sigma_{\epsilon}^2),
\label{eq:multiple_glmm}
\end{aligned}
\end{equation}
where $\textrm{I}_{\textrm{group}}$ is a indicator matrix for groups and $\otimes$ is the Kronecker product.

>>>>

\newcommand{\plmefour}{phyloglmm/lme4}
\newcommand{\pgTMB}{phyloglmm/glmmTMB}
\newcommand{\phyvar}{\sigma} %% phylogenetic var: allow sigma/Sigma switch
\newcommand{\phyv}[1]{\ensuremath{\phyvar^2_{\mathrm{phy}_\mathrm{#1}}}}
\newcommand{\phys}[1]{\ensuremath{\phyvar_{\mathrm{phy}_\mathrm{#1}}}}
\newcommand{\bphys}[1]{\ensuremath{{\mathbf b}_{\mathrm{phy}_{\mathrm{#1}}}}}
\newcommand{\bsps}[1]{\ensuremath{{\mathbf b}_{\mathrm{sp}_{\mathrm{#1}}}}}
\newcommand{\bgrp}[1]{{{\mathbf b}_{\mathrm{#1}}}}
% \pagenumbering{gobble}
\linenumbers

@@

We extend the simulation model by adding multiple groups where each group has one observation per species. 
The multi-group model is a generalization of multiple-site models used in community ecology to model phylogenetic attraction and repulsion \citep{helmus2007separating}. 
The full multi-group model is as follows:

\begin{equation}
\begin{aligned}
\bY & = \dist(g^{-1}(\bmu), \phi) \\
\bmu & = (\beta_0 + \bphys{int} + \bsps{int} + \bgrp{group}) + (\beta_1 + \bphys{slope} + \bsps{slope}) \bm{t} + \bgrp{sp:group} + \besp \\
(\bphys{int}, \bphys{slope}) & \sim \textrm{MVN} \left( 0, \begin{bmatrix}
\phyv{int} & \phys{int-slope} \\ 
\phys{int-slope} & \phyv{slope}
\end{bmatrix}
\right) \\
(\bf{b_{\mathrm{sp_{int}}}, b_{\mathrm{sp_{slope}}}}) & \sim \textrm{MVN} \left( 0, \begin{bmatrix}
\sigma^2_{\mathrm{sp_{int}}} & \sigma_{\mathrm{sp_{int-slope}}} \\ 
\sigma_{\mathrm{sp_{int-slope}}} & \sigma^2_{\mathrm{sp_{slope}}}
\end{bmatrix}
\right) \\
\bf{b_{\mathrm{group}}} & \sim \textrm{MVN} ( 0 , \sigma^2_{\mathrm{group}}) \\
\bf{b_{\mathrm{sp:group}}} & \sim \textrm{MVN} (0, \mathrm{I}_{\mathrm{group}} \otimes \sigma^2_{\mathrm{phy}}) \\
\besp & \sim \textrm{N} ( 0 , \sigma_{\epsilon}^2),
\label{eq:multiple_glmm}
\end{aligned}
\end{equation}
where $\textrm{I}_{\textrm{group}}$ is a indicator matrix for groups and $\otimes$ is the Kronecker product.
